Non-word repetition in Yélî Dnye

### 2020-12-15 manuscript.Rmd
minor edits to methods text

This study was approved as part of a larger research effort by Marisa Casillas, with the research protocol having been approved by the IRB committee entitled The Ethics Committee of the Faculty of Social Sciences (Ethiek Commissie van de faculteit der Sociale Wetenschappen; ECSW) of Radboud University, Nijmegen, The Netherlands.  The approval for work outside of the Netherlands came as an amendment to the general research line on language development research (original request: ECSW2017-3001-474 Manko-Rowland; amendment: ECSW-2018-041). Participation was voluntary, with children being invited to come and participate. Regardless of whether they did any of the task, children were provided with a snack as compensation. Children who came up to participate but then refused were nonetheless provided with the snack.

@@

A total `r dim(demo)[1]` children were tested, from `r length(levels(factor(demo$familyID)))` of families, in four hamlets. Some children could not be included for the following reasons: refused participation or failed to repeat items presented over headphones even after coaching (N=`r sum(demo$included=="didn\xeat repeat")`), spoke too softly to allow offline coding (N=`r sum(demo$included=="spoke too softly")`). In addition, `r sum(demo$Age >12 | demo$included=="too old",na.rm=T)` teenagers were tested to put younger children at ease; their data is not included in analyses below. The remaining `r dim(inc)[1]` children (`r sum(inc$Sex=="G")` girls) were aged `r round(mean(inc$age.rounded),2)` years (range `r round(min(inc$age.rounded),2)`-`r round(max(inc$age.rounded),2)` years). Among these, there were `r sum(inc$monolingual=="yes", na.rm=T)` children exposed only to Yélî in the home, `r sum(inc$monolingual=="no", na.rm=T)` children who were also exposed to another language in the home, and `r sum(is.na(inc$monolingual))` for whom this information is missing. Maternal years of education averaged `r round(mean(inc$matEd,na.rm=T),2)` years (range `r round(min(inc$matEd,na.rm=T),2)`-`r round(max(inc$matEd,na.rm=T),2)` years; `r sum(is.na(inc$matEd))` children had this information missing).[^1] In terms of birth order, `r birthorder.tab[1]` were first borns, `r birthorder.tab[2]` second,  `r birthorder.tab[3]` third,  `r birthorder.tab[4]` forth,  `r birthorder.tab[5]` fifth, `r birthorder.tab[6]` sixth, and  `r birthorder.tab[7]` did not have this information.

[^1]: Precise years of education can be hard to count. This is particularly the case for schooling beyond grade 8, which takes place off-island, which can be due to an irregularity in ability to pay school fees, school opening times, familial obligations, and more. We asked parents to report their highest completed level of education and inferred the approximate number of years that entails under ideal conditions, which results in many esimates around typical "completion" times in between.

>>>>

This study was approved as part of a larger research effort by the second author. The line of research was evaluated by the Radboud University Faculty of Social Sciences Ethics Committee (Ethiek Commissie van de faculteit der Sociale Wetenschappen; ECSW) in Nijmegen, The Netherlands (original request: ECSW2017-3001-474 Manko-Rowland; amendment: ECSW-2018-041). Participation was voluntary; children were invited to participate following indication of approval from an adult caregiver. Regardless of whether they completed the task, children were given a small snack as compensation. Children who showed initial interest but then decided not to participate were also given the snack.

@@

We tested a total of `r dim(demo)[1]` children from `r length(levels(factor(demo$familyID)))` families spread across four hamlets. We excluded test sessions from analysis for the following reasons: refused participation or failed to repeat items presented over headphones even after coaching (N=`r sum(demo$included=="didn\xeat repeat")`), spoke too softly to allow offline coding (N=`r sum(demo$included=="spoke too softly")`), or were 13 years old or older (N=`r sum(demo$Age >12 | demo$included=="too old",na.rm=T)`; we tested these teenagers to put younger children at ease). The remaining `r dim(inc)[1]` children (`r sum(inc$Sex=="G")` girls) were aged `r round(mean(inc$age.rounded),2)` years (range `r round(min(inc$age.rounded),2)`-`r round(max(inc$age.rounded),2)` years). Among these children were `r sum(inc$monolingual=="yes", na.rm=T)` only exposed to Yélî Dnye at home and `r sum(inc$monolingual=="no", na.rm=T)` children exposed to one or more other languages at home. Maternal years of education averaged `r round(mean(inc$matEd,na.rm=T),2)` years (range `r round(min(inc$matEd,na.rm=T),2)`-`r round(max(inc$matEd,na.rm=T),2)` years).[^1] In terms of birth order, `r birthorder.tab[[1]]` were first borns, `r birthorder.tab[[2]]` second,  `r birthorder.tab[[3]]` third,  `r birthorder.tab[[4]]` forth,  `r birthorder.tab[[5]]` fifth, and `r birthorder.tab[[6]]` sixth.

[^1]: Years of education is difficult to precisely pinpoint in this context. Schooling beyond grade 8 takes place off-island and enrollment is tied to an ability to pay that year's school fees, seasonal interruptions to school opening times, familial obligations, and more. We asked parents to report their highest completed level of education. We then record the number of years entailed by having completed that level under ideal conditions. This method results in many esimates around typical "graduation" times (e.g., graduating from primary education before starting secondary education).
